---
title: "a3_task3_devin_ngo"
author: "Devin Ngo"
date: "10/25/2021"
output: html_document
---

# Setup
```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
```

# Read in Urchins Data 

```{r}
# Read in Data
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

# Section 1: Size Comparisons of Purple and Red Urchins

## Data Visualization 

```{r}
ggplot(data = urchins,
       aes(x = size)) +
  geom_histogram(bins = 10, fill = "sienna2", color = "black") +
  facet_wrap(~common_name) +
  labs(x = "Size",
       y = "Count",
       title = "Purple and Red Urchin Size Count")
```

## Descriptive Data Table
Create a finalized summary table that contains the following statistics for purple and red urchins (still ignoring site as a variable, but including species as a grouping variable): mean, median, standard deviation, and sample size. Use kableExtra (or another package for tables in R) to finalize your table for your knitted document. Add a table caption above your table. 

```{r}
urchins_summary<- urchins %>% 
  group_by(common_name) %>% 
  summarize (mean_size = mean(size),
             median_size = median(size),
             sd_size = sd(size),
             n = n()) %>% 
  kable(col.names = c("Species",
                      "Mean Size",
                      "Median Size",
                      "Standard Deviation Size",
                      "Sample Size")) %>% 
  kable_styling(full_width = FALSE)

urchins_summary
```

