---
title: "Naples Reef Purple Urchins (2010)"
author: "Devin Ngo"
date: "10/25/2021"
output: html_document
---
# Introduction

This data explores the size frequency distribution of red and purple sea urchins recorded by Santa Barbara Coastal Long Term Ecological Research program divers at long-term kelp removal experiment sites. For this report, we will specifically be looking at data for purple urchins in the Naples Reef in 2010. 

**Data Citation:** Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

# Setup
```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# Attach Packages
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```

# Read in Urchins Data
```{r}
# Read in data 

urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

# Wrangling Data 
Wrangling the urchins data to only include Naples Reef observations for purple urchins observed in 2010. 
```{r}
purple_urchins_2010 <- urchins %>% 
  filter(common_name %in% c("Purple Urchin")) %>% 
  mutate(date_class = mdy(date)) %>% 
  mutate(year = year(date_class)) %>% 
  filter(year %in% c("2010")) %>% 
  filter(site %in% c("NAPL"))
```

# Data Visualization 
Exploratory faceted histogram for the size distribution of purple urchins at Naples in 2010, separated by treatment (Control and Annual), to assess normality.
```{r}
ggplot(data = purple_urchins_2010, 
       aes(x = size)) + 
  geom_histogram(aes(bins = 12, fill = treatment)) +
  facet_wrap(~treatment)
```

Exploratory QQ-plots for the size distribution of purple urchins at Naples in 2010, separated by treatment (Control and Annual), to assess normality.
```{r}
ggplot(data = purple_urchins_2010) + 
  geom_qq(aes(sample = size, color = treatment)) +
  facet_wrap(~treatment)
```


```{r}
purple_urchins_summary <- purple_urchins_2010 %>% 
  group_by(treatment) %>% 
  summarize (mean_size = round(mean(size),3),
             median_size = round(median(size),3),
             sd_size = round(sd(size),3),
             n = n()) %>% 
  kable(col.names = c("Treatment",
                      "Mean Size",
                      "Median Size",
                      "Standard Deviation of Size",
                      "Sample Size (n)")) %>% 
  kable_styling(full_width = FALSE)

purple_urchins_summary
```


